Hierarchical wrinkling in a confined permeable biogel

### 2015-02-16 fig_plis/article_plis.tex
typo Thomas

To form wrinkled patterns such as the one observed in figure~\ref{fig:dynamics}, ones needs to take advantage of the pH sensitivity of caseins to control the route to gelation, to fine tune the gel properties and the boundary conditions.

@@

Boundary conditions is the final parameter that is essential to form wrinkled patterns. To turn off the adsorption of the caseins on a surface, we coat it with polymer brushes (see Methods), effectively creating non adhesive boundary conditions for the gel. The evolution of $\xi$ when sticky boundaries are on reflects the inner constrains of the gel: $\xi$ increases, the stands of the network become thinner and the gel wants to retract; then $\xi$ decreases, the stands of the network swell and the gel wants to expend. When the casein adhesion to the upper boundary is turned off, the gel still not wrinkles because it remains attached to the bottom wall, but it detaches from the top wall. In this situation we can monitor $\xi$, Figure~\ref{fig:acidification}d. $\xi$ remains mostly constant but the volume occupied by the gel varies, as will be reported below.

@@

The determination of the characteristic length $\lambda$ of a wrinkling pattern is an old problem. Large deformations of thin sheets is governed by F\"{o}ppl-von K\'arm\'an equations~\cite{landau1986}, essentially impossible to solve in analytical form. However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern and isolate the physically meaningful ingredients~\cite{Cerda2003}. An elastic film buckles due to excess area with respect to its boundaries. Wrinkling is a buckling hindered by a substrate: the mismatch of elastic properties between the film and the substrate selects the preferred wrinkling wavelength~\cite{Gough1940, Bijlaard1946}. $\lambda$ is then set by the competition between the flexural modulus of the film $B=E/(1-\nu^2)$ ($E$ Young modulus, $\nu$ Poisson ratio) and a transverse load $\sigma_\perp$. For small deflections of  amplitude $A$, $\lambda$ scales as~\cite{Vella2009,Kolinski2009}: 
%
\begin{equation}
\lambda \sim \left( A h^3 \frac{B}{\sigma_{\perp}} \right)^{1/4}.
\label{eq:lstar}
\end{equation}

@@

The fraction of free casein proteins of a \SI{100}{\milli\litre}, 4\%w sodium caseinate dispersion is obtained either by titration by HCl 1 molar or during the action of 4\%w GDL. In both cases, \SI{0.4}{\milli\litre} of the \SI{100}{\milli\litre} dispersion is sampled regularly as a function of the pH, and centrifuged in eppendorf tubes at 14500 rpm for \SI{3}{\minute}. The supernatant is diluted and analysed with a UV absorption spectrometer (Ocean Optics). According to~\cite{Roefs1986}, we deduced the fraction of free casein proteins from the adsorption spectrum:
%
\begin{align}
x_{free} &= \frac{A(\text{pH})}{A(\text{pH}=7)},\\
\text{with }A(\text{pH}) &= A_{280}-1.706 A_{320},
\end{align}
%
where $A_{280}$ is the absorption peak due to aromatic amino acids in the casein molecules plus a turbidity background and $A_{320}$ is only due to turbidity.

>>>>

To form wrinkled patterns such as the one observed in figure~\ref{fig:dynamics}, ones needs to take advantage of the pH sensitivity of caseins, to control the route to gelation, to fine tune the gel properties and the boundary conditions.

@@

Boundary conditions is the final parameter that is essential to form wrinkled patterns. To turn off the adsorption of the caseins on a surface, we coat the surface with polymer brushes (see Methods), effectively creating non adhesive boundary conditions for the gel. The evolution of $\xi$ when sticky boundaries are on reflects the inner constrains of the gel: $\xi$ increases, the stands of the network become thinner and the gel wants to retract; then $\xi$ decreases, the stands of the network swell and the gel wants to expend. When the casein adhesion to the upper boundary is turned off, the gel still not wrinkles because it remains attached to the bottom wall, but it detaches from the top wall. In this situation we can monitor $\xi$, Figure~\ref{fig:acidification}d. $\xi$ remains mostly constant but the volume occupied by the gel varies, as will be reported below.

@@

The determination of the characteristic length $\lambda$ of a wrinkling pattern is an old problem. Large deformations of thin sheets is governed by F\"{o}ppl-von K\'arm\'an equations~\cite{landau1986}, which are essentially impossible to solve in analytical form. However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern and isolate the physically meaningful ingredients~\cite{Cerda2003}. An elastic film buckles due to excess area with respect to its boundaries. Wrinkling is a buckling hindered by a substrate: the mismatch of elastic properties between the film and the substrate selects the preferred wrinkling wavelength~\cite{Gough1940, Bijlaard1946}. $\lambda$ is then set by the competition between the flexural modulus of the film $B=E/(1-\nu^2)$ ($E$ Young modulus, $\nu$ Poisson ratio) and a transverse load $\sigma_\perp$. For small deflections of  amplitude $A$, $\lambda$ scales as~\cite{Vella2009,Kolinski2009}: 
%
\begin{equation}
\lambda \sim \left( A h^3 \frac{B}{\sigma_{\perp}} \right)^{1/4}.
\label{eq:lstar}
\end{equation}

@@

The fraction of free casein proteins of a \SI{100}{\milli\litre}, 4\%w sodium caseinate dispersion is obtained either by titration by HCl 1 molar or during the action of 4\%w GDL. In both cases, \SI{0.4}{\milli\litre} of the \SI{100}{\milli\litre} dispersion is sampled regularly as a function of the pH, and centrifuged (centrifuge 5424, Eppendorf) in Eppendorf tubes at 14500 rpm for \SI{3}{\minute}. The supernatant is diluted and analysed with a UV absorption spectrometer (USB4000, Ocean Optics). According to~\cite{Roefs1986}, we deduced the fraction of free casein proteins from the adsorption spectrum:
%
\begin{align}
x_{free} &= \frac{A(\text{pH})}{A(\text{pH}=7)},\\
\text{with }A(\text{pH}) &= A_{280}-1.706 A_{320},
\end{align}
%
where $A_{280}$ is the absorption peak due to aromatic amino acids in the casein molecules plus a turbidity background and $A_{320}$ is only due to turbidity.
